Understanding the impacts of video-guided activities on parent-child interaction

### 2021-01-08 paper/manuscript.Rmd
minor revisions

author: 
  - name          : "George Kachergis"
    affiliation   : "1"
    address       : "Stanford, CA 94305 USA"
    email         : "kachergis@stanford.edu"
  - name          : "Emily Hembacher"
    affiliation   : "1"
  - name          : "Veronica Cristiano"
    affiliation   : "2"
  - name          : "Hanwen Vivian Zhang"
    affiliation   : "3"
  - name          : "Michael C. Frank"
    corresponding : yes
    address       : "Stanford, CA 94305 USA"
    email         : "mcfrank@stanford.edu"
    affiliation   : "1"

@@

abstract: |
  Early parenting practices play an important role in shaping children’s future outcomes. 
  In particular, high-quality early interactions can facilitate language learning and school performance. 
  The rise of phone-based parenting applications ("apps") could deliver low-cost interventions on parenting style to a wide variety of populations, especially the parents of very young children, who are often difficult to reach in other ways. 
  Yet little is known about the effects of communicating to parents through app-based interventions. 
  In two studies (one preregistered), we showed parents short videos depicting age-appropriate parent-child activities from a parenting app. 
  We found that after watching the video, parents spoke more and made more bids for joint attention, as compared with controls who watched no video (experiment 1) or a science video (experiment 2). 
  These results suggest that activity videos can lead to positive changes in parent engagement, providing support for the use of such videos in parenting interventions.
keywords          : "language development; parent intervention; childhood development; joint attention; lexical diversity"
wordcount         : "4954"
bibliography      : ["library.bib"]
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
header-includes:
  - \usepackage{float}
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=4, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=T, 
                      message=F, sanitize = T)
```

@@

The content provided by digital parenting interventions runs the gamut from general parenting messages and facts from child development research to specific advice and suggested activities. 
A growing body of evidence suggests that these digital interventions can be effective across a range of cultures, income levels, and children's ages [for a review, see @Breitenstein2014].
For example, in contrast to a face-to-face parent training intervention, a tablet-based version saw significantly higher session completion rates (51% attendance vs. 85% module completion) and comparable or larger effect sizes on parents' and children's (aged 2 to 5 years) behavior [@Breitenstein2016].
Often, however, the theory of change presupposed by such interventions is relatively vague.
Both within and outside the realm of academic interventions, messages to parents of young children often seek to provide knowledge about some aspect of development (e.g., early language), often in tandem with a suggestion regarding activities.
Such messages are assumed to inform parents' choice of behaviors, spurring them to engage in some target activity, which is assumed to be more stimulating than what parents would have done otherwise.

@@

In Experiment 1, we invited parents of 6- to 24-month-old infants visiting the Children's Discovery Museum in San Jose to complete video-guided activities from a commerical parenting app that delivers digital parenting advice in the form of short videos.
Parents were randomly assigned to one of two conditions: parents in the _Activity Video_ condition watched a video from the app (matched to their child's age), and then performed the activity with their child using the props from the video. 
Parents in the _No Video_ condition did not watch an activity video, but were given a set of the same age-appropriate props and asked to play with their infants as they normally would at home.

@@

### Participants. 
`r nrow(e1ld)` infants (F = `r e1_gender['F']`, M = `r e1_gender['M']`) aged 6-24 months (19 6-11.9 month-olds: 9 in the control condition, 10 in Activity Video; 20 12-17.9 month-olds: 11 control, 9 Activity Video; 21 18-24 month-olds: 10 control, 11 Activity Video) and their parents participated in a museum in northern California. 
We included infants who were exposed to English at least 50 percent of the time (n = 58) or who were exposed less but whose participating parent reported that they primarily speak English with their child at home (n = 2). 
62% of participants (n = 37) had been exposed to two or more languages, as indicated by their parent. 
Parents identified their children as White (n = 25), Asian (n = 11), African American/Black (n = 2), Biracial (n = 12), other (n = 5), or declined to state (n = 5). 
Fifteen parents reported that their child was of Hispanic origin. 
Parents tended to be highly-educated, with reports of highest level of education ranging from completed high school (n = `r e1_parent_ed[1]`), some college (n = `r e1_parent_ed[2]`), four-year college (n = `r e1_parent_ed[3]`), some graduate school (n = `r e1_parent_ed[4]`), to completed graduate school (n = `r e1_parent_ed[5]`).

### Materials.
Stimuli included activity videos from a commercial parenting application. 
The videos were designed to show activities to parents that they could perform with their child in order to foster cognitive and physical development, and were targeted to the child's age and level of development. 
In each video, an adult and child perform the activity (e.g., sorting toys according to size) while a narrator explains the activity and its purpose. 
We selected two videos for each of three age groups in our sample (6-11.9 months, 12-17.9 months, 18-23.94 months).
Participants were also given a set of toys corresponding to those in the video that they watched so that they could complete the activity.[^1]

@@

### Procedure.
After providing informed consent, parents in the Activity Video condition watched the assigned activity video on a laptop with headphones.
To ensure that parents could give the video their full attention, the experimenter played with the infant with a set of toys (different from the experimental props used in the study) while the video was being played. 
Immediately following the video, each parent-child dyad was provided with the props to complete the video-guided activity that the parent had viewed. 
The toys were placed on a large foam play mat, and parents were instructed to sit on the mat with their child and re-create the activity they had viewed for a period of approximately three minutes.[^2]
In the No Video condition, after informed consent parents were told to play with their child as they would at home with the provided props for a period of three minutes. 
They were not given any additional instructions about how to use the props.

@@

### Joint Attention Coding Procedure.
The video of each session was manually coded for episodes of joint attention (JA) using the Datavyu software [@datavyu]. 
The video taken at floor level was coded by default, but the other video was referred to if the participants were occluded or if there was technical difficulty with the first camera. 
Each session's video was coded for episodes of coordinated JA, episodes of passive JA, and parental bids for JA. 
Parental bids for JA were defined as any attempt to initiate joint attention (i.e labeling, pointing, or otherwise drawing attention to an object) that did not result in passive or coordinated JA. 
If more than 3 seconds elapsed between bids, they were coded as separate attempts. 
An episode of joint attention was considered _passive_ if both participants visually focused on an object for 3 or more seconds but the child did not acknowledge the parent. 
If either participant looked away from the object for less than 3 seconds and then returned to the same object it was considered part of the same episode of joint attention. 
A joint attention episode was considered _coordinated_ if both participants visually focused on an object for 3 or more seconds and at some point in the interaction the child indicated awareness of interaction with some overt behavior toward the parent such as looking at their face, gesturing, vocalizing, or turn-taking. 
Full details of our guidelines for coding joint attention are available in SI.

@@

### Participants. 
`r nrow(e2ld)` infants (F = `r e2_gender['F']`, M = `r e2_gender['M']`) aged 12-24 months 
(`r e2_12to18["Activity Video"]` 12-17.9 month-olds in the Activity Video condition; 
`r e2_12to18["Science Video"]` 12-17.9 month-olds in the Science Video condition; 
`r e2_18to24["Activity Video"]` 18-24 month-olds in the Activity Video condition;
`r e2_18to24["Science Video"]` 18-24 month-olds in the Science Video condition) and their parents participated in the same museum as Experiment 1. 
We included infants who were exposed to English at least 75 percent of the time or who were exposed less but whose participating parent reported that they primarily speak English with their child at home. 
Forty-nine percent of participants (n = 41) had been exposed to two or more languages as indicated by their parent. 
Parents identified their children as White (n = 39), Asian (n = 20), African American/Black (n = 1), Biracial (n = 9), other (n = 7), or declined to state (n = 8). 
Sixteen parents reported their child was of Hispanic origin. 
Parents tended to be highly-educated, with reports of highest level of education ranging from some college (n = `r e2_parent_ed["3"]`), four-year college (n = `r e2_parent_ed["4"]`), some graduate school (n = `r e2_parent_ed["5"]`), to completed graduate school (n = `r e2_parent_ed["6"]`) or declined to state (n = 13).

### Materials.

The design of Experiment 2 was similar to that of Experiment 1, except that instead of seeing no video in the control condition, parents instead watched a video that was generally related to child development research, but did not give any specific instructions about how to interact with infants or children. 
This condition was included to control for the possibility that differences in language output and joint attention in Experiment 1 could be due to simply cueing parents to think about infants' learning and cognitive development. 
The videos presented in the Control Video condition were media clips (available on YouTube) of developmental psychologists explaining their research interleaved with footage of infants or toddlers engaged in developmental research studies. 
Thus, the content of the videos superficially matched those in the Activity Video condition, but did not suggest any particular activities.
The videos were trimmed to approximately match the average video length in the Activity Video condition (close to 90 s).
Details of the videos used in the Activity Video conditions are in the Appendix.

### Procedure. 
The procedure for Experiment 2 matched that of Experiment 1, except that parents in the Control Video condition watched a control video before the play session. 
Consistent with the No-Video control condition in Experiment 1, parents in the Control Video condition were told to play with their child as they would at home, and were not given additional instructions.
The coding procedure also matched that of Experiment 1. 
To establish reliability a second coder independently coded 25 of the 84 videos, approximately equally distributed across ages. 
The two coders had a reliability of ICC = 0.81 with 95% confidence interval (CI) = [0.62,0.91] for number of parent bids for JA; ICC = 0.74 with 95% CI = [0.48, 0.88] for number of passive JA episodes; ICC = 0.80 with 95% CI = [0.61, 0.91] for number of coordinated JA episodes; ICC = 0.72 with 95% CI = [0.44, 0.86] for time spent (secs/min) in passive JA episodes, and ICC = 0.88 with 95% CI = [0.75, 0.94] for time spent in coordinated JA episodes.

>>>>

author: 
  - name          : "George Kachergis"
    corresponding: yes
    affiliation   : "1"
    address       : "Stanford, CA 94305 USA"
    email         : "kachergis@stanford.edu"
  - name          : "Emily Hembacher"
    affiliation   : "1"
  - name          : "Veronica Cristiano"
    affiliation   : "2"
  - name          : "Hanwen Vivian Zhang"
    affiliation   : "3"
  - name          : "Michael C. Frank"
    address       : "Stanford, CA 94305 USA"
    email         : "mcfrank@stanford.edu"
    affiliation   : "1"

authornote: |
  The data that support the findings of this study are openly available in OSF at http://doi.org/10.17605/OSF.IO/2BPDF

@@

abstract: |
  Early parenting practices play an important role in shaping children’s future outcomes. 
  In particular, high-quality early interactions can facilitate language learning and school performance. 
  The rise of phone-based parenting applications ("apps") could deliver low-cost interventions on parenting style to a wide variety of populations, especially the parents of very young children, who are often difficult to reach in other ways. 
  Yet little is known about the effects of communicating to parents through app-based interventions. 
  In two studies (one preregistered), we showed parents short videos depicting age-appropriate parent-child activities from a parenting app. 
  We found that after watching the video, parents spoke more and made more bids for joint attention, as compared with controls who watched no video (experiment 1) or a science video (experiment 2). 
  These results suggest that activity videos can lead to positive changes in parent engagement, providing support for the use of such videos in parenting interventions.
keywords          : "language development; parent intervention; childhood development; joint attention; lexical diversity"
wordcount         : "5066"
bibliography      : ["library.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
header-includes:
  - \usepackage{float}
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r global_options, include=FALSE}
# The data that support the findings of this study are openly available in OSF at https://osf.io/2bpdf/?view_only=d4b230deba2a4f91bdb2aba29088d982
#  The data that support the findings of this study are openly available in OSF at http://doi.org/10.17605/OSF.IO/2BPDF

knitr::opts_chunk$set(fig.width=4, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=T, 
                      message=F, sanitize = T)
```

@@

The content provided by digital parenting interventions runs the gamut from general parenting messages and facts from child development research to specific advice, coaching, and suggested activities. 
A growing body of evidence suggests that these digital interventions can be effective across a range of cultures, income levels, and children's ages [for a review, see @Breitenstein2014].
For example, in contrast to a face-to-face parent training intervention, a tablet-based version saw significantly higher session completion rates (51% attendance vs. 85% module completion) and comparable or larger effect sizes on parents' and children's (aged 2 to 5 years) behavior [@Breitenstein2016].
Often, however, the theory of change presupposed by such interventions is relatively vague.
Both within and outside the realm of academic interventions, messages to parents of young children often seek to provide knowledge about some aspect of development (e.g., early language), often in tandem with a suggestion regarding activities.
Such messages are assumed to inform parents' choice of behaviors, spurring them to engage in some target activity, which is assumed to be more stimulating than what parents would have done otherwise.

@@

In Experiment 1, we invited parents of 6- to 24-month-old infants visiting the Children's Discovery Museum in San Jose to complete video-guided activities from a commercial parenting app that delivers digital parenting advice in the form of short videos.
Parents were randomly assigned to one of two conditions: parents in the _Activity Video_ condition watched a video from the app (matched to their child's age), and then performed the activity with their child using the props from the video. 
Parents in the _No Video_ condition did not watch an activity video, but were given a set of the same age-appropriate props and asked to play with their infants as they normally would at home.

@@

### Participants
`r nrow(e1ld)` infants (F = `r e1_gender['F']`, M = `r e1_gender['M']`) aged 6-24 months (19 6-11.9 month-olds: 9 in the control condition, 10 in Activity Video; 20 12-17.9 month-olds: 11 control, 9 Activity Video; 21 18-24 month-olds: 10 control, 11 Activity Video) and their parents participated in a museum in northern California. 
We included infants who were exposed to English more than 50 percent of the time (n = 58) or who were exposed less but whose participating parent reported that they primarily speak English with their child at home (n = 2). 
62% of participants (n = 37) had been exposed to two or more languages, as indicated by their parent. 
Parents identified their children as White (n = 25), Asian (n = 11), African American/Black (n = 2), Biracial (n = 12), other (n = 5), or declined to state (n = 5). 
Fifteen parents reported that their child was of Hispanic origin. 
Parents tended to be highly-educated, with reports of highest level of education ranging from completed high school (n = `r e1_parent_ed[1]`), some college (n = `r e1_parent_ed[2]`), four-year college (n = `r e1_parent_ed[3]`), some graduate school (n = `r e1_parent_ed[4]`), to completed graduate school (n = `r e1_parent_ed[5]`).

### Materials
Stimuli included activity videos from a commercial parenting application. 
The videos were designed to show activities to parents that they could perform with their child in order to foster cognitive and physical development, and were targeted to the child's age and level of development. 
In each video, an adult and child perform the activity (e.g., sorting toys according to size) while a narrator explains the activity and its purpose. 
We selected two videos for each of three age groups in our sample (6-11.9 months, 12-17.9 months, 18-23.94 months).
Participants were also given a set of toys corresponding to those in the video that they watched so that they could complete the activity.[^1]

@@

### Procedure
After providing informed consent, parents in the Activity Video condition watched the assigned activity video on a laptop with headphones.
To ensure that parents could give the video their full attention, the experimenter played with the infant with a set of toys (different from the experimental props used in the study) while the video was being played. 
Immediately following the video, each parent-child dyad was provided with the props to complete the video-guided activity that the parent had viewed. 
The toys were placed on a large foam play mat, and parents were instructed to sit on the mat with their child and re-create the activity they had viewed for a period of approximately three minutes.[^2]
In the No Video condition, after informed consent parents were told to play with their child as they would at home with the provided props for a period of three minutes. 
They were not given any additional instructions about how to use the props.

@@

### Joint Attention Coding Procedure
The video of each session was manually coded for episodes of joint attention (JA) using the Datavyu software [@datavyu], and coders were blind to condition. 
The video taken at floor level was coded by default, but the other video was referred to if the participants were occluded or if there was technical difficulty with the first camera. 
Each session's video was coded for episodes of coordinated JA, episodes of passive JA, and parental bids for JA. 
Parental bids for JA were defined as any attempt to initiate joint attention (i.e labeling, pointing, or otherwise drawing attention to an object) that did not result in passive or coordinated JA. 
If more than 3 seconds elapsed between bids, they were coded as separate attempts. 
An episode of joint attention was considered _passive_ if both participants visually focused on an object for 3 or more seconds but the child did not acknowledge the parent. 
If either participant looked away from the object for less than 3 seconds and then returned to the same object it was considered part of the same episode of joint attention. 
A joint attention episode was considered _coordinated_ if both participants visually focused on an object for 3 or more seconds and at some point in the interaction the child indicated awareness of interaction with some overt behavior toward the parent such as looking at their face, gesturing, vocalizing, or turn-taking. 
Full details of our guidelines for coding joint attention are available in SI.

@@

### Participants
`r nrow(e2ld)` infants (F = `r e2_gender['F']`, M = `r e2_gender['M']`) aged 12-24 months 
(`r e2_12to18["Activity Video"]` 12-17.9 month-olds in the Activity Video condition; 
`r e2_12to18["Science Video"]` 12-17.9 month-olds in the Science Video condition; 
`r e2_18to24["Activity Video"]` 18-24 month-olds in the Activity Video condition;
`r e2_18to24["Science Video"]` 18-24 month-olds in the Science Video condition) and their parents participated in the same museum as Experiment 1. 
We included infants who were exposed to English more than 50 percent of the time or who were exposed less but whose participating parent reported that they primarily speak English with their child at home. 
Forty-nine percent of participants (n = 41) had been exposed to two or more languages as indicated by their parent. 
Parents identified their children as White (n = 39), Asian (n = 20), African American/Black (n = 1), Biracial (n = 9), other (n = 7), or declined to state (n = 8). 
Sixteen parents reported their child was of Hispanic origin. 
Parents tended to be highly-educated, with reports of highest level of education ranging from some college (n = `r e2_parent_ed["3"]`), four-year college (n = `r e2_parent_ed["4"]`), some graduate school (n = `r e2_parent_ed["5"]`), to completed graduate school (n = `r e2_parent_ed["6"]`) or declined to state (n = 13).

### Materials
The design of Experiment 2 was similar to that of Experiment 1, except that instead of seeing no video in the control condition, parents instead watched a video that was generally related to child development research, but did not give any specific instructions about how to interact with infants or children. 
This condition was included to control for the possibility that differences in language output and joint attention in Experiment 1 could be due to simply cueing parents to think about infants' learning and cognitive development. 
The videos presented in the Control Video condition were media clips (available on YouTube) of developmental psychologists explaining their research interleaved with footage of infants or toddlers engaged in developmental research studies. 
Thus, the content of the videos superficially matched those in the Activity Video condition, but did not suggest any particular activities.
The videos were trimmed to approximately match the average video length in the Activity Video condition (close to 90 s).
Details of the videos used in the Activity Video conditions are in the Appendix.

### Procedure
The procedure for Experiment 2 matched that of Experiment 1, except that parents in the Control Video condition watched a control video before the play session. 
Consistent with the No-Video control condition in Experiment 1, parents in the Control Video condition were told to play with their child as they would at home, and were not given additional instructions.
The coding procedure also matched that of Experiment 1. 
To establish reliability a second coder independently coded 25 of the 84 videos, approximately equally distributed across ages. 
The two coders had a reliability of ICC = 0.81 with 95% confidence interval (CI) = [0.62,0.91] for number of parent bids for JA; ICC = 0.74 with 95% CI = [0.48, 0.88] for number of passive JA episodes; ICC = 0.80 with 95% CI = [0.61, 0.91] for number of coordinated JA episodes; ICC = 0.72 with 95% CI = [0.44, 0.86] for time spent (secs/min) in passive JA episodes, and ICC = 0.88 with 95% CI = [0.75, 0.94] for time spent in coordinated JA episodes.
